A Study of Immersion Factors Using an Augmented Reality Simulation

### 2009-06-24 Paper/paper.tex

Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.  The ability to see through solid objects has applications in various fields ranging from engineering and architecture to military or search-and-rescue operations.  Unfortunately, implementing x-ray vision is not a trivial task.  AR systems have cost issules, including those pertaining to level-of-immersion.  For example, what is the ideal augmented field of view (FOV) for a given scenario?  Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where tracking is temporarily lost.

@@

We can more explicitly detect the lost state by applying a threshold $\tau$ to the data.  When the error rises above the threshold, we consider the participant lost.  TODO: how to determine this threshold??

One metric which uses this threshold is the ``time to failure,'' the time until a lost state is encountered.  However a participant may reach the lost state sooner or later depending on the movement of the avatars.  We also considered a measurement we called ``time tracking,'' which is the amount of time spent in the tracking state.  This metric seems to most clearly represent how well a participant performed.  Trials with either more lost periods or longer lost periods will result in a lower total time tracking.

% count number of times lost

% time tracking is overview metric -- captures both number of times lost and length of lost periods
% number of times lost is piece of overview -- number of times lost.

\begin{figure}[ht!]
	\centering
	\includegraphics[width=3in]{figures/tophatscreenshot.png}
	\caption{Sample user-view of AR simulation.  \emph{Top Hat} is visible (non-occluded) in the doorway.  The augmented view overlays transparent red rectangles on the characters, indicating their location while occluded.}
\end{figure}

@@

% should we include data from older generation??

% look at analysis from mean.

% time tracking: fov, deadlen significant.  interaction is weird....

% num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.

>>>>

Augmented Reality systems may be used to simulate x-ray vision \cite{1383060}.  The ability to see through solid objects has potential applications in various fields ranging from construction \cite{Webster96augmentedreality}, medicine \cite{azuma95survey}, military \cite{Livingston02anaugmented} and search-and-rescue operations \cite{1528424}.  Unfortunately, implementing x-ray vision is not a trivial task.  AR systems have cost issules, including those pertaining to level-of-immersion.  For example, what is the ideal augmented field of view (FOV) for a given scenario?  Other issues relate to tracking the objects in the x-ray view, such as GPS dropout errors \cite{4079263}, where tracking is temporarily lost.

@@

We can more explicitly detect the participant's state by applying a threshold $\tau$ to the data.  The threshold specifies the maximum angular error that still represents successful tracking of the target. 
%When the error rises above the threshold, we consider the participant lost.
We use a threshold of one quarter of the total horizontal field of view of the HMD, $\tau = 12$ degrees.
% TODO: how to motivate this threshold?

Using this threshold we measure a participant's ``time to failure,'' the time until the a lost state is encountered.  However, this may not be a very descriptive measure, since a participant may reach the lost state sooner or later depending on the movement of the avatars, which varies between trials.

We also consider a measurement we call ``time tracking,'' which is the amount of time spent in the tracking state.  This metric seems to most generally represent how well a participant performed.  Trials with either more lost periods or longer lost periods will result in a lower total time tracking.

For a more specific measure of performance, we count the number of times a participant reaches the lost state in a trial.  We only record a lost state when the error stays above the threshold $\tau$ for a minimum of 0.5 seconds.
% TODO: how to motivate this threshold?

%time tracking is overview metric -- captures both number of times lost and length of lost periods

%number of times lost is piece of overview -- number of times lost.

@@

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_deadlen.pdf}
%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods.}
%\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_fov.pdf}
%	\caption{Box plot of the number of times the participant becomes lost versus the field of view of the augmentations.}
%\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_interaction.pdf}
%	\caption{Interaction plot of the effect of field of view and tracking failure on the number of times the participant becomes lost.}
%\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=3in]{figures/tt_deadlen.pdf}
	\caption{\label{fig:tt_deadlen}Box plot of the amount of time that the target is near the crosshairs versus the length of tracker failure periods.}
\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/tt_fov.pdf}
%	\caption{Box plot of time the participant is not lost versus the field of view of the augmentations.}
%\end{figure}

%\begin{figure}[t]
%	\centering
%	\includegraphics[width=3in]{figures/numtimes_extra_deadlen.pdf}
%	\caption{Box plot of the number of times the participant is lost versus the length of the tracking failure periods, for the medium field of view condition, including the half second trial.}
%\end{figure}

% TODO: should we include data from older generation??

% TODO: look at analysis from mean.

We performed two-way {ANOVA} for the time tracking metric as an overall analysis of the effects of our two factors.  Table \ref{fig:anova} gives the results of this analysis.  Both the failure length and the field of view had a very significant effect on performance using this metric.  
% TODO: interaction?

% TODO: num times lost: fov not significant, deadlen significant, interaction -- talk about what interaction means.

Figure \ref{fig:tt_deadlen} shows a box plot of time tracking versus failure length.  Performance decreases between zero and one seconds, but seems to level off between one and two.  We ran a post-hoc analysis to determine which conditions were significantly different.  Using Tukey's HSD, we found that conditions one and two were significantly different than zero (p<??), but not significantly different from each other (p<??).  This suggests that we may be seeing a thresholding effect above one second tracking failures.

This suggests that we have reached a threshold at one second.  We had seven participants complete an extra three trials with the medium field of view and tracking failure periods of 0.5 seconds.  Figure 8 shows the data.  There wasn't a significant difference between the zero length and half second length conditions.  For future work, we could find the threshold between 0.5 and 1.

%use multiple comparison of means to determine which pairs are significant -- difference between length of one and length of two is not significant.

\begin{figure}[htb]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
%\multicolumn{2}{|c|}{Selection Stroke Thresholds} \\
Metric & Factor & F value & Pr \\
\hline
Time Tracking & Failure Length & 20.41 & $<0.001$\\
Time Tracking & Field of View & 13.549 & $<0.001$\\
Number of Times Lost & Failure Length & 61.816 & $<0.001$\\
\hline
\end{tabular}
 \caption{\label{fig:anova}Factors which have a significant effect.}
\end{figure}
